---
title: "Assess Forecast of Opportunity's Probability of Success"
# mainfont: DejaVu Sans
output:
  pdf_document:
    # toc: true
    df_print: kable
    toc_depth: 4
    # number_sections: true
---

<!-- Data to explore: -->

```{r setup,  cache = TRUE, echo = FALSE, message = FALSE}
wd <- getwd()
setwd("../R_OPPO_PROB")
source("../R_OPPO_PROB/_probGiveInLast.R", local = TRUE)
gd <- "2018-10-01"
data <- probGiveInLast(given_date = gd)
```

```{r, transform, cache = FALSE, echo = FALSE, message = FALSE}
require(data.table)
require(tidyr)
require(dplyr)

prob <- data[, .(Opportunity_ID,
                 `Predicted Probability`,
                 `Probability %`,
                 Winning_Probability,
                 # IsClosed,
                 # IsWon,
                 # now_IsClosed,
                 now_IsWon)]

p4plot <- gather(prob, key = "prob_type", value = "probability",
                 `Predicted Probability`,
                 "Winning_Probability",
                 "Probability %")

setDT(p4plot)

p4plot[, prob_type := case_when(prob_type %chin% "Predicted Probability" ~ "Machine Learning",
                                prob_type %chin% "Probability %" ~ "Sales Force CRM",
                                prob_type %chin% "Winning_Probability" ~ "Project Manager",
                                TRUE ~ "Other")]

```


Based on `r gd` Open Opportunities relative to it's actual outcomes\


  
```{r, echo = FALSE, fig.height = 3.5, message = FALSE}

require(ggplot2)
require(ggthemes)
require(data.table)
require(dplyr)

ggplot(data = p4plot, aes(x = prob_type, y = probability, fill = prob_type)) +
        geom_boxplot(alpha = 0.99, outlier.shape = 20) +
        facet_wrap(~ case_when(now_IsWon == 0 ~ "LOST",
                               now_IsWon == 1 ~ "WON",
                               TRUE ~ "NA")) +
        theme_bw() +
        # ggthemes::theme_hc()
        # xlab(NULL) +
        # theme_bw()
        theme(axis.title.x = element_blank(),
              axis.text.x  = element_blank(),
              axis.ticks.x = element_blank(),
              legend.title = element_blank(),
              legend.position = "top",
              legend.margin = margin(0,0,0,0),
              legend.box.margin = margin(-7,-7,-7,-7)) +
        scale_fill_brewer(type = "qual", palette = 7, direction = 1) +
                labs(
             # subtitle = "Based on 2018-10-01 Open Opportunities") +
             caption = paste0("Total Opportunity IDs: ",
             uniqueN(p4plot$Opportunity_ID)))
# scale_fill_brewer()

        # geom_jitter(alpha = .02, size = 4, width = 0.01, height = 0.2)
```
*Legend annotation:*\
> `Machine Learning` - is my Machine Learning model stored in SQL as **`Predicted Probability`** field.\
> `Sales Force CRM` - SalesForce Automated Probabilities stored in SQL as **`Probability %`** field.\
> `Project Manager` - Account Manager's best guess stored in SQL as  **`Winning Probability`** field.\

  
\   

```{r, message = FALSE, echo = FALSE}
require(data.table)
require(dplyr)
require(knitr)
require(kableExtra)
p4plot <- as.data.table(p4plot)

pcast <- p4plot[, .("Mean Probabilty" = round(mean(probability), 2),
                    "Median Probabilty" = round(median(probability), 2)),
       by = .("PROBABILITY SOURCE" = prob_type,
              "now_IsWon" = case_when(now_IsWon == 0 ~ "LOST",
                                      now_IsWon == 1 ~ "WON",
                                      TRUE ~ "OTHER"))]

```


  
\  
  


Median Summary Statistics:
```{r, message = FALSE, echo = FALSE}
require(data.table)
require(dplyr)
as_tibble(dcast(pcast, `PROBABILITY SOURCE` ~ now_IsWon,
                value.var = "Median Probabilty")) %>% 
                  kable() %>%
  kable_styling(full_width = T, bootstrap_options = "striped") %>% 
        column_spec(1, bold = T)
```


  
\  
  
    
    
    
\  


Mean Summary Statistics:

```{r, message = FALSE, echo = FALSE}
as_tibble(dcast(pcast, `PROBABILITY SOURCE` ~ now_IsWon, value.var = "Mean Probabilty")) %>% 
          kable() %>%
  kable_styling(full_width = T, bootstrap_options = "striped") %>% 
        column_spec(1, bold = T)
```



```{r, echo = FALSE, message = FALSE, fig.height = 7}
require(ggplot2)
require(data.table)
require(ggthemes)
ggplot(data = p4plot, aes(x = probability,
                          y = case_when(now_IsWon == 0 ~ "LOST",
                                        now_IsWon == 1 ~ "WON",
                                        TRUE ~ "NA"),
                          color = prob_type)) +
        geom_jitter(alpha = .10, width = .04, height = 0.15, size = 3, color = "black") +
        # geom_boxplot() +
        facet_wrap(~ prob_type , ncol = 1) +
        ylab("Actual Outcome") +
        xlab("Probability Forecast") +
        scale_color_brewer(type = "qual", palette = 7, direction = 1) +
        theme_bw() +
        theme(legend.position = "none")

```



